---
title: "project-performance"
author: "er637474"
date: "April 16, 2018"
output: html_document
---

```{r setup, include=FALSE}
library(ProjectTemplate)
load.project(Eduardo Rosillo csx415)
knitr::opts_chunk$set(echo = FALSE)
```

#### MTech Energy Solutions: Initial Analysis

Within this document there is an initial analysis of the total usage (kilowatts) and monthly cost (dollars US) over an annual period (calculated per month of a total year) of the San Francisco hotel property. This analysis is evaluated under the standard metric referred to in the Formal Problem Statement. Also included is a inferenial test, a Pearson's product-moment correlation, to evaluate the how well the analysis performs.

```{r, include=FALSE}
require(ggplot2)
require(Hmisc)
projectdata <- pd.csv <- read.csv("C:/Users/eduardo/Desktop/projectdata.csv")
str(projectdata)
knitr::opts_chunk$set(echo = FALSE)

```


```{r, include=FALSE}
1:12
c(1,10,11)
projectdata[1:12,c(1,10,11)]
knitr::opts_chunk$set(echo = FALSE)

```
Below is a basic summary of the initial analysis of the Total Usage (top line) and Cost monthly (bottom line) per month over an annual period.

```{r, include=TRUE}

with(projectdata, summary(Total_Usage))
with(projectdata, summary(Cost_Monthly))
knitr::opts_chunk$set(echo = FALSE)

```

#### S.F.H. property Data Visualized:


```{r , include=TRUE}
ggplot(data=projectdata, aes(x=Month, y=Total_Usage, group = 1)) + 
  geom_line(color = "black", linetype = "dashed", size =1.5) + 
  geom_point(color = "black", size = 4, shape=21, fill = "black") +
  scale_x_discrete(limits = month.name)


ggplot(data=projectdata, aes(x=Month, y=Cost_Monthly, group = 1))+ 
  geom_line(color = "black", linetype = "dashed", size =1.5) + 
  geom_point(color = "black", size = 4, shape=21, fill = "black") +
  scale_x_discrete(limits = month.name)

knitr::opts_chunk$set(echo = FALSE)

```

Using Pearson Product Moment Correlation we will measeure how well the energy usage and demand is related to the monthly operational cost of the S.F.H. property. With a cor value = 0.899 the PPMC method reveals a strong correlation between the energy usage/demand of the property and the monthly costs.

PPMC is not able to tell the difference between dependent variables and independent variables. To put this in perspective of the project; this means that while an increase in energy usage per month could increase the monthly cost, spending more money on energy usage would increase demand in energy per month. This makes no sense as common practice is to pay for the energy demand after it has been delivered at time of billing. This is a significant downside to this correlation test.
```{r, include=TRUE}
pd.quant <- projectdata[c(10,11)]
cor(pd.quant)
cor.test(pd.quant$Total_Usage, pd.quant$Cost_Monthly)
rcorr(as.matrix(pd.quant)) 

knitr::opts_chunk$set(echo = FALSE)

```

In order to furhter improve the analysis, it is recommended that a model be developed to predict the monthly costs based on the energy usage per month and a root-mean-square deviation test used to evaluate the accuracy of said model. 